Allocation scores, WIS, and CRPS via decision theory

### 2023-08-23 alloscore_manuscript/methods_apg.tex
reorganizing in terms of scoring rules rather than procedure steps

Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} made up of pre-specified actions, states and consequences and rating the utility of that solution.  That is, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.  It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.  But as scores accumulate, we can use their distributions across forecasting instances to make inferences as to whether one forecasting method is better or worse than another in a decision theoretic sense.

@@

\section{Step 1: Defining decision problems and loss functions.}

We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$. The resource has a unit-cost of $C$, and we assign a loss of $L$ to the consequence of a sick individual not having access to the resource. This specifies a loss function  
\begin{align}
s_Q(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
\end{align}
where $u_+ = \max(u,0)$.  We use the subscript $Q$ because it will be shown below that this loss function yields the quantile score for the the probability level $\alpha = 1 - C/L$.  Loss functions of this general type are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.  (We ignore for now any discreteness of the resource.) The consequence set $\mathcal{C}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$. We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need. This would imply a Bayes act $x^F$ that is identically 0, telling us nothing about the value of $F$.

This decision problem may however be complicated by uncertainty at time $t_0$ about the unit-cost of the resource. We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.  Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified consequences of the decision under unrealized costs. If we wanted to commit to a \emph{locality} principle for prediction where losses and scores only depend on predictive densities in a neighborhood of the relevant observation, this attempt would falter, and we would need to frame our methods around the log score and it's higher order relatives. But we do not see that as an appropriate principle in public health resourse allocation where  
probabilistic forecasts can be valuable by pushing decision makers in the "right direction" even when they assign low probability to observed outcomes. We instead adopt a ``sensitivity-to-distance'' principle from which piece-wise linear loss functions and the QS can be shown to naturally emerge. It is this principle that makes the loss under alternative costs seem relevant. \apg{These are obviously incomplete thoughts but I feel like there is something here that I've been grasping for for a lng time.}

A simple way to take account of the consequences of actions under counterfactual costs is to use the mean of all losses under certainty about the cost:
\begin{align}
s_{\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}C_ix_i + L(x_i - y)_+ \quad (C_i < L \text{ for all } i)
\end{align}
where we here use $x$ to express the decision vector $(x_1,\ldots, x_m)$. Or if there are prior beliefs about the probabilities $p_i$ of the $C_i$, we could use a weighted sum
\begin{align}
s_{w\mathrm{WIS}}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
\end{align}
As for $Q$, the $WIS$ and $wWIS$ subscripts indicate that, for particular values of the $\{C_i\}$ and $L$, this loss function produces the WIS or a version that weights intervals (i.e. the average of QS's for $C_i$ and $L-C_i$) with non-standard weights (combining $p_i$ and $C_i/L$).

We can also consider $C$ a continuous random variable and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} which (with some mathematical care) converts the loss into an integral
\begin{align}
s_{\mathrm{CRPS}}(x,y) = \frac{1}{L}\int_{0}^{L} (cx(c) + L(x(c) - y)_+) dc
\end{align}
over all relevant costs $c \in [0,L]$, or a weighted version
\begin{align}
s_{w\mathrm{CRPS}}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}
\end{align}
that uses a prior density $f_C$ on the possible costs. These loss functions will yield the traditional CRPS and the wieghted version of \cite{gneiting2011weightedScoringRules}.

@@

We can also expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.  This includes the motivating scenario for our work, the simultaneous procurement of resources across several locations (U.S. states and territories) during a pandemic. And again, summing the losses  \eqref{eqn:quantile_loss} over the locations (for a particular $C$ and $L$) creates a loss function that is reactive to the overall procurement decision. Let the locations be indexed by $l \in 1,\ldots,N$.  The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_i\}, i = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
\begin{align}
\mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N\\
\mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
\end{align}
And for these "vector-action" problems we can define \emph{total} versions of our loss functions
\begin{align}
s_{t\mathrm{Q}}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
s_{t\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}C_i x_{l,i} + L(x_{l,i} - y_l)_+ \\
s_{tw\mathrm{WIS}}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
s_{t\mathrm{CRPS}}(x,y) &= \frac{1}{L} \sum_{l=1}^N\int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) dc \\
s_{tw\mathrm{CRPS}}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
\end{align}
A key point for our presentation will be that the same $C$, set $\{C_i\}$, probabilities $\{p_i\}$, and density $f_C$ are used for all locations simultaneously.  We believe that our methods could be adapted to variations across locations (which could be important in practice), but we leave this generalization for later work.

@@

Consider again the simultaneous resource procurement decision a public health worker must make in attempting to mitigate a pandemic across multiple locations. 
We now add a constraint $K$ on the total amount of the resource available for allocation across all locations. This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_l \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost). Loss functions with the basic $Cx + L(y-x)_{+}$ structure we have used up until now continue to generate meaningful scoring rules with respect to the AP following our procedure.  But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_l$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
\begin{align}
\mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
\end{align}
In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
This isolates -- at least formally -- the explicit contribution of the procurement part, $x$, of the decision to the second term which we call the \emph{allocation loss function}:
\begin{align}
s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
\end{align}
It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.

\section{Step 2: Derive the Bayes act for a forecast $F$}

I think we should not hide the formula 
\begin{align}
\dby{}{x_{il}} E_F[s_{.}(x,y)] &= E_{F}[\dby{}{x_{il}} s_{.}(x,y)] = C - LE_{F_l}[\mathbf{1}(y_l > x_{il})] = C + L(F_l(x_{il}) - 1) \\
\implies x_{il}^{F_l} &= F_l^{-1}(1 - C/L) \\
&\text{ or for CRPS} \\
x_{l}^{F_l}(c) &= F_l^{-1}(1 - c/L)
\end{align}

@@

\subsection{Solving the Allocation Problem}

As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K, \lambda^{\star})$ in the region $\{\mathbf{x} \geq 0, \lambda \geq 0\}$ for the Lagrangian 
\begin{align}
\mathcal{L}(\mathbf{x},\lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
\end{align}
That is, optimality at $x^F$ requires that
\begin{align}
0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^F, z_K, \lambda) &= L E_F \left[\frac{d}{dx_l}(Y_l - x_l^F)_{+}\right] + \lambda \\
& = L E_F \left[-\mathbf{1}\{Y_i \geq x_i^F\}\right] + \lambda \\
& = L(F_i(x_l^F)-1) + \lambda \label{eqn:x_stationary}
\end{align}
for any $x_l^F > 0$;
\begin{align}
0 = \frac{\partial}{\partial z_K} \mathcal{L}(x^F, z_K, \lambda) &= \lambda - C
\end{align}
if $z_K>0$; and
\begin{align}
0 = \del{}{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
& = \sum_{\{i | x_i^F > 0\}} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
\end{align}
These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see ...). We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.  This search begins with the interval $\lambda \in [0,1]$ and 
checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.  If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.  Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for 
$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.  This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an 
approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.

>>>>

Broadly speaking, we score a probabilistic forecast $F$ by identifying it with a solution to a \emph{decision problem} and rating the utility of that solution.  The term ``decision problem'' refers to a pre-specified set of permissable actions, possible future states, and resulting consequences of actions under various states.  Accordingly, $F$ will receive a "better" or "worse" score in a particular instance when the consequence is better or worse of a rational decision maker using $F$ (and essentially only $F$) to select an action in the face of uncertainty about a future state.  It should be noted immediately that a better or worse score in a single instance may tell us almost nothing about the intrinsic value of the forecasting method that produced $F$.  But as scores accumulate, we can use their distributions across forecasting instances to make inferences as to whether one forecasting method is better or worse than another in a decision theoretic sense.

@@

\section{The quantile score}

We begin by imagining a public health worker trying to decide at time $t_0$ what quantity $x$ to procure of a protective resource such as beds, staff, or medicine in anticipation of $y$ new infections that will be observed (and create resource demand) at time $t_1>t_0$. The resource has a unit-cost of $C$, and we assign a loss of $L$ to the consequence of a sick individual not having access to the resource. This specifies a loss function  
\begin{align}
s_{Q,C}(x,y) = Cx + L(y-x)_+ \label{eqn:quantile_loss}
\end{align}
where $u_+ = \max(u,0)$.  Loss functions of this general form are referred to as \emph{piece-wise linear}, distinguishing them from smooth non-linear loss functions such as squared error.
The underlying decision problem here consists of action and outcome sets $\mathcal{X},\mathcal{Y}$ both equal to $\mathbb{R}_+$, the non-negative reals.  (We ignore for now any discreteness of the resource.) The consequence set $\mathcal{C}$ is abstractly a set of over-expenditure regrets and numbers of sick individuals with unmet resource need, but the loss function $s$ maps $\mathcal{C}$ into $\mathbb{R}_+$. We also from now on make the assumption that $C < L$ since otherwise -- at least according to this loss function -- there is no ``cost-effectiveness'' in trying to prevent even a single case of unmet need. This would imply a Bayes act $x^F$ that is identically 0, telling us nothing about the value of $F$.

The Bayes act for $C \in (0,L)$ is the solution of the first order equation 
\begin{align}
0 = \dby{}{x} E_F[s_{Q,C}(x,y)] &= E_{F}[\dby{}{x}s_{Q,C}(x,y)] \\
&= C + LE_{F}[\dby{}{x}(y-x)_+] \\
&= C - LE_{F}[\mathbf{1}\{y > x\}] \\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
that is, the quantile $x^F = F_l^{-1}(1 - C/L) = q_{F,\tau}$ for the probability level $\tau = 1-C/L$. (That the critical point $q_{F,\tau}$ is actually a minimum follows from \eqref{eqn:q_deriv} being non-positive to its left and non-negtive to its right, and assuming as we do for now that $F$ is smooth and strictly increasing at $q_{F,\tau}$, it is a unique minimum.) The quantile solves the decision problem with respect to the forecast $F$ by being the action which best balances expected cost against expected loss among all actions available to the decision maker.

Finally, we evaluate the scoring rule $S_Q$ on $F$ when outcome $y$ realizes by measuring the loss that occurs when the action $q_{F,\tau}$ is taken:
\begin{align}
S_Q(F,y) = s_Q(q_{F,\tau}, y) = Cq_{F,\tau} + L(y- q_{F,\tau})_{+}.
\end{align}

\section{WIS and CRPS as quantile scores under cost uncertainty}

The decision problem leading to the QS may however be complicated by uncertainty at time $t_0$ about the cost $C$ of the resource. We take such uncertainty as requiring the worker to commit in advance to procurement levels $x_i$ for a range of potential costs $C_i < L$ where $i \in 1, \ldots, m$.  Even though only one of the $x_i$ will end up needing to be procured, there does not seem to be any intrinsically decision theoretic principle by which our loss function should ignore counter-factual but exactly specified consequences of the decision under unrealized costs. If we wanted to commit to a \emph{locality} principle for prediction where losses and scores only depend on predictive densities in a neighborhood of the relevant observation, this attempt would falter, and we would need to frame our methods around the log score and it's higher order relatives. But we do not see that as an appropriate principle in public health resourse allocation where  
probabilistic forecasts can be valuable by pushing decision makers in the "right direction" even when they assign low probability to observed outcomes. We instead adopt a ``sensitivity-to-distance'' principle from which piece-wise linear loss functions and the QS can be shown to naturally emerge. It is this principle that makes the loss under alternative costs seem relevant. \apg{These are obviously incomplete thoughts but I feel like there is something here that I've been grasping for for a lng time.}

We take into account the consequences of actions under a set $\mathcal{C} = \{C_i\}$ of possible costs by expanding the loss function to be a weighted mean of quantile losses under certainty about the cost:
\begin{align}
s_{Q,\mathcal{C}, p_C}(x,y) = \frac{1}{m}\sum_{i=1}^{m}p_i(C_ix_i + L(x_i - y)_+). \label{eqn:sum_CL_loss}
\end{align}
Here, $p_C = \{p_i\}, i = 1,\ldots,m$ is a prior distribution on $\mathcal{C}$ and $x = (x_1,\ldots,x_m)$ is the decision vector corresponding to all possible costs.  $p_C$ would be provided by the decision maker (or perhaps a separate expert), and we emphasize that we are considering it here as part of the loss function $s_{Q,\mathcal{C}, p_C}$ that would be formulated in an effort to adapt forecast evaluation to the specifics of a public health decision problem. In practice, $p_C$ has usually been taken to be uniform, which we will show for a particular $\mathcal{C}$ (given $L$) leads to the WIS familiar in ID forecasting hubs. This might be interpreted as deferring to an uninformative prior.

We can also consider $C$ a continuous random variable with a prior density $f_C$ and accommodate this by letting $m \to \infty$ in \eqref{eqn:sum_CL_loss} 
which (with some mathematical care) converts the weighted mean into an integral
\begin{align}
s_{Q,f_C}(x,y) = \int_{0}^{L} (cx(c) + L(x(c) - y)_+) f_C(c)dc{}.
\end{align}

@@

The Bayes acts for losses $s_{Q,\mathcal{C}, p_C}$ and $s_{Q,f_C}$ therefore correspond not just to the vanishing of the derivative with respect to a single
decision variable $x$ but to the vanishing of the partial derivatives with respect to all $m$ decision variables $\{x_i\}$ in the first case or the infinitely many variables $\{x(c)\}$ in the second case.  But since the losses are sums of single varianble losses, passing to solutions of these higher dominsional equations is a mere formality of the viewing the set of Bayes acts, i.e. quantiles, for costs $c \in \mathcal{C}$ or $(0,L)$ as a vector 
$(x^F(1),\ldots,x^F(m))=(q_{F,\tau_1},\ldots,q_{F,\tau_m})$ or a function $x(c) = q_{F,\tau(c)}$.

Similarly, given an outcome $y$, evaluation of the scoring rules $s_{Q,\mathcal{C}, p_C}$ and $s_{Q,f_C}$ on the forecast $F$ amounts to vectorizing the
formula for $S_Q(F,y)$. That is,
\begin{align}
S_{Q,\mathcal{C}, p_C}(F,y) &= \frac{1}{m}\sum_{i=1}^{m}p_i(C_i q_{F,\tau_i} + L(q_{F,\tau_i} - y)_+) \\
S_{Q, f_C}(F,y) &= \int_{0}^{L}(cq_{F,\tau(c)} + L(q_{F,\tau(c)} - y)_+) f_C(c)dc.
\end{align}

Again, the hierarchy we have introduced here describing WIS and CRPS as parallels to the quantile score for higher dimensional action sets is in itself a mere formality employing the linearity of differentiation with respect to the expectation operator $E_{C}$ for $C$ considered as a random variable. But as we move now to decision problems with multiple outcomes, it will turn out to provide a key shift in perspective that allows the scoring rules we define in terms of allocation under scarcity to be seen as a natural and non-trivial extension of the hierarchy.

\section{Aggregating QS, WIS, and CRPS across dimensions}

We next expand our decision problem to include multiple actions that will all, in fact (not counterfactually), have to be taken.  This includes the motivating scenario for our work, the simultaneous procurement of resources across several locations (U.S. states and territories) during a pandemic. And again, summing the losses  \eqref{eqn:quantile_loss} over the locations (for a particular $C$ and $L$) creates a loss function that is reactive to the overall procurement decision. Let the locations be indexed by $l \in 1,\ldots,N$.  The action sets for the decision problem corresponding to a single certain cost $C$, a discrete cost distribution over $\{C_i\}, i = 1,\ldots,m$, and a continuous cost distribution $f_C$ with support on $[0,L]$ are now
\begin{align}
\mathcal{X}_{cert} &= \mathbb{R}_{+}^N \\
\mathcal{X}_{disc} &= (\mathbb{R}_{+}^m)^N = \{x(i):\{1,\ldots,m\} \to \mathbb{R}_{+}^N\}\\
\mathcal{X}_{cont} &= \{x(c):[0,\infty) \to \mathbb{R}_{+}^N\}.
\end{align}
And for these ``vector-action'' problems we can define \emph{total} versions of our loss functions
\begin{align}
s_{tQ}(x,y) &= \sum_{l=1}^N Cx_l + L(y_l-x_l)_+ \\
s_{tQ, \mathcal{C}, p_C}(x,y) &= \frac{1}{m}\sum_{l=1}^N \sum_{i=1}^{m}p_i(C_i x_{l,i} + L(x_{l,i} - y_l)_+)\\
s_{tQ, f_C}(x,y) &= \sum_{l=1}^N \int_{0}^{L} (cx_l(c) + L(x_l(c) - y_l)_+) f_C(c)dc.
\end{align}
A key point for our presentation will be that the same $C$, set $\{C_i\}$, probabilities $\{p_i\}$, and density $f_C$ are used for all locations simultaneously.  We believe that our methods could be adapted to variations across locations (which could be important in practice), but we leave this generalization for later work.

\section{The allocation score}

@@

Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_i, i \in 1,\ldots,m$) a public health worker must make in 
attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.  But now we ask the worker to extend their decision to cover the contingency that the total amount of the resource $\sum x_l(C)$ that they specify for the realized cost $C$ turns out to not be available.  This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures. In either case, we assume that these factors are undetermined at time $t_0$ when decisions must be made and model them with a positive random variable $K$ giving the total amount of resource that will be available
at time $t_1$.  This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resourse is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose? The action set for this new component is 
\begin{align}
\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid \sum_{l=1}^{N} x_l = k\}.
\end{align}
Note that because this decision is only relevant when $\sum x_l(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)

In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left(x, y\right) = Ck + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
We see then that the cost $C$ will not effect the Bayes act for this component of the decision problem motivating the definition of a new single loss function (component), the \emph{allocation loss function}
\begin{align}
s_{A}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+},
\end{align}
we use for all $\mathcal{X}_k$.

\subsubsection*{Old allocation intro}
We now add a constraint $K$ on the total amount of the resource available for allocation across all locations. This presents the worker with a new decision problem we call the \emph{allocation problem} (AP or $\mathrm{AP}(K)$) with a smaller action set $\mathcal{X}_K = \{x \in \mathbb{R}_{+}^N | \sum x_l \leq K\} \subset \mathcal{X}_{cert}\}$ from which to choose (and analogously restricted version of the sets $\mathcal{X}_{disc}$ and $\mathcal{X}_{cont}$ for the decision problems with uncertain cost). Loss functions with the basic $Cx + L(y-x)_{+}$ structure we have used up until now continue to generate meaningful scoring rules with respect to the AP following our procedure.  But with a view toward developing a scoring rule that accommodates uncertainty about both $C$ and $K$ in a non-redundant (and more interpretable and computationally feasible) way, we augment the action set of the AP to include, in addition to the procurement levels $x_l$ for all locations, a ``reserve'' quantity $z_K$ defined as the available amount of resource that the worker decides not to allocate anywhere.
In optimization theory quantities defined in this manner are known as \emph{slack} variables, and are introduced in order to re-express contraint inequalities as equalities.  In our case, we get
\begin{align}
\mathcal{X}_{K} = \{(x,z) \in \mathbb{R}_{+}^{N+1} \mid \sum_{l=1}^{N}x_l + z = K \}.
\end{align}
In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left((x,z), y\right) = C(K-z) + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
This isolates -- at least formally -- the explicit contribution of the procurement part, $x$, of the decision to the second term which we call the \emph{allocation loss function}:
\begin{align}
s_{A,K}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+}.
\end{align}
It will turn out, as explained in the following sections, that for a given forecast $F$, a scoring rule derived with respect to the AP from $s_{tQ}$ under uncertainty about $C$ -- that is, from $s_{tWIS}$ or $s_{tCRPS}$ -- will have the same values for all outcomes $y$ as a scoring rule that is a weighted average or integration of scoring rules derived from allocation losses $s_{A,K}$ with respect the AP's for a range of $K$'s.

@@

\section{Solving the Allocation Problem}

As the solution to a constrained optimization problem, the optimal allocation $x^F$ according to a forecast $F$ can be described formally as the first component of a stationary point $(x^F, z_K^F, \lambda^{\star})$ in the region $\{x, z \geq 0, \lambda \in \mathbb{R}\}$ for the Lagrangian 
\begin{align}
\mathcal{L}(x, z_K, \lambda; K, C, L) = L\sum_{l=1}^{N} E_{F_l}[(Y_l - x_l)_{+}] + C(K-z_K) + \lambda\left(z_K + \sum_{l=1}^{N} x_l - K\right).
\end{align}
That is, optimality at $x^F$ requires that
\begin{align}
0 = \frac{\partial}{\partial x_l} \mathcal{L}(x^F, z_K, \lambda) &= L E_F \left[\frac{d}{dx_l}(Y_l - x_l^F)_{+}\right] + \lambda \\
& = L E_F \left[-\mathbf{1}\{Y_l \geq x_l^F\}\right] + \lambda \\
& = L(F_l(x_l^F)-1) + \lambda \label{eqn:x_stationary}
\end{align}
for any $x_l^F > 0$;
\begin{align}
0 = \frac{\partial}{\partial z_K} \mathcal{L}(x^F, z_K, \lambda) &= \lambda - C
\end{align}
if $z_K>0$; and
\begin{align}
0 = \del{}{\lambda} \mathcal{L}(x^F, z_K^F, \lambda^{\star}) &= \sum_{l=1}^{N} x_l^F + z_K^F - K\\
& = \sum_{\{l | x_l^F > 0\}} F_{l}^{-1}(1-\lambda^{\star}) + z_K^F - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
\end{align}
These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
when they are, a solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can only be obtained analytically in special cases (but see ...). We therefore typically calculate allocation scores using approximately optimal allocations found via a binary search for $\lambda^{\star}$.  This search begins with the interval $\lambda \in [0,1]$ and 
checks at each step the sign of the inequality that results from substituting the interval midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.  If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.  Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for 
$1-\lambda_{\tau}$ do not make full use of our resource and we set $\lambda_{\tau}$ as the upper endpoint of the $\tau + 1$ search interval.  This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$ yielding an 
approximation $\lambda^{\star, \varepsilon}$. When quantiles $F_i^{-1}(1-\lambda^{\star, \varepsilon})$ are singly defined and the forecast densities $f_i$
not too close to zero at these quantiles, the discrepancy $\Delta(\lambda^{\star, \varepsilon})  = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star, \varepsilon}) - K$ is of order $\varepsilon_{\lambda}$ and we can take $F_{i}^{-1}(1-\lambda^{\star, \varepsilon})$ as approximations to $x_i^F$.
